Behavior Driven Development in Haskell

### 2012-06-02 paper.tex
Work on reusable specs

Haddock before version 2.8 had markup syntax for
code blocks, but no markup syntax for evidence.
As mentioned before, we implemented support for interactive examples ({\tt >>>}) in 2010.
It is available in Haddock version 2.8 or later.
We also added support for markup of properties ({\tt prop>}) which
will be available in a future release of Haddock.
Both interactive examples and properties in Haddock comments are
included in the generated documentation.

@@

Reusable spec is an idea to scrap boilerplate of evidence.
For instance, without reusable spec,
each developer of an instance of a type class should
express evidence for its laws.
Since instances of a type class share API,
we can avoid such boilerplate.
The following is an example to implement
evidence for Monoid laws:

@@

\noindent This can be used to document and test arbitrary Monoid
instances.
Here is an example how to use it for lists:

\begin{quote}
\small
\begin{verbatim}
describe "List as a Monoid" $
  shouldSatisfyMonoidLaws (undefined :: [Int])
\end{verbatim}
\end{quote}

\noindent We have already implemented this experimentally.
We should ensure the practicality of reusable spec and
prepare reusable spec to cover variety of type classes.

@@

\begin{quote}
    TBD: state that ``A library for reusable pieces of specification would be
    useful.''
\begin{itemize}
    \item
        In Section~\ref{sec:reusable-specs}, we showed how to reuse
        pieces of specification.
    \item
        It would be possible and useful to write a library that
        provides reusable pieces of specification, for many of
        Haskell's standard classes.
        (similar in purpose to
        {\tt
        checkers}\footnote{http://hackage.haskell.org/package/checkers})
\end{itemize}
\end{quote}

>>>>

Haddock before version 2.8 had markup syntax for
code blocks, but not for evidence.
We implemented support for interactive examples ({\tt >>>}) in
Haddock 2.8.
We also added support for markup of properties ({\tt prop>}) which
will be available in a future release.
Both interactive examples and properties in Haddock comments are
included in the generated documentation.

@@

RSpec provides an explicit mechanism to facilitate reuse of pieces of
specification.~\cite{rspec}
In Haskell we get reusability for free, it does not require any
explicit support from Hspec.

\begin{quote}
  TBD: not sure, if this adds anything

This not only shows that functional programming can deliver what OO
has always claimed, but it is also awfully useful.
\end{quote}

Here is a specification that documents that a list is a monoid:

\begin{quote}
\small
\begin{verbatim}
spec = do
  describe "List as a Monoid" $ do
    describe "mempty" $ do
      it "is a left identity" $ property $ \x ->
        mempty <> x == x

      it "is a right identity" $ property $ \x ->
        x <> mempty == x

    describe "<>" $ do
      it "is associative" $ property $ \x y z ->
        (x <> y) <> z == x <> (y <> z)
\end{verbatim}
\end{quote}

\noindent And it is possible to to make this specification reusable,
by abstracting over the type:

@@

\noindent This can be used to document and test arbitrary Monoid
instances, the specification from above would become:

\begin{quote}
\small
\begin{verbatim}
spec = do
  describe "List as a Monoid" $ do
    shouldSatisfyMonoidLaws (undefined :: [Int])
\end{verbatim}
\end{quote}

@@

\begin{quote}
    TBD: state that ``A library for reusable pieces of specification would be
    useful.''
\begin{itemize}
    \item
        In Section~\ref{sec:reusable-specs}, we showed how to reuse
        pieces of specification.
    \item
        It would be possible and useful to write a library that
        provides reusable pieces of specification, for many of
        Haskell's standard classes.
        (similar in purpose to
        {\tt
        checkers}\footnote{http://hackage.haskell.org/package/checkers})
      \item copied here from Section~\ref{sec:reuseable-pecs}:
        We should ensure the practicality of reusable spec and
        prepare reusable spec to cover variety of type classes.

\end{itemize}
\end{quote}
